Urbanization and Public Transportation of Los Angeles

### 2023-11-27 Paper.Rmd
Better exploratory analysis

---
title: "Urbanization and Public Transportation of Los Angeles"
author: "Kainoa Kanter, Julian Jacobson, Aidan Allen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    df_print: paged
  pdf_document:
    extra_dependencies: amsmath
abstract: This paper will examine how Los Angeles's demand for public transportation
  has been affected the last few years, utilizing derivatives to determine the rate
  of change in public transportation usage and studying the elasticity of demand for
  transportation options based on pricing, availability, convenience, and modes.
geometry: margin=1in
---

@@

#### Elasticity of Demand Analysis:

-   Analyze the responsiveness of public transportation demand to changes in price (fare changes), availability (service hours, frequency), convenience (route changes, wait times), and modes (bus vs. rail).

-   Employ econometric modeling to estimate the price elasticity of demand, cross-elasticity, and income elasticity, using regression analysis where ridership

@@

```{r rate-of-change-plot, fig.cap="Line plot of rate of change in public transportation ridership over time"}
ridership_long <- gather(ridership_data, key = "Type", value = "RateOfChange", -Year)

@@

-   Transit ridership data: Los Angeles Metro "L.A. METRO TRANSIT RIDERSHIP UP 10 PERCENT, SETS POST-PANDEMIC RECORD", Patrick Chandler

>>>>

---
title: "Urbanization and Public Transportation of Los Angeles"
author: "Kainoa Kanter, Julian Jacobson, Aidan Allen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    df_print: paged
  pdf_document:
    extra_dependencies: amsmath
abstract: This paper will examine how Los Angeles's demand for public transportation
  has been affected the last few years, utilizing derivatives to determine the rate
  of change in public transportation usage.
geometry: margin=1in
---

@@

@@

As seen in the graph below, ridership data for both the Metro Bus system and Metro Light Rail between 2019 and 2023 has been heavily impacted by the 2020 pandemic. The bus ridership has had more of a sharp decline from 2019 to 2020 and bounce-back shortly, while the rail was slowing down from 2019 - 2021, with not a lot of recovery.

```{r rate-of-change-plot, fig.cap="Line plot of rate of change in public transportation ridership over time", fig.height=4}
ridership_long <- gather(ridership_data, key = "Type", value = "RateOfChange", -Year)

@@

-   Transit ridership data: Los Angeles Metro "L.A. METRO TRANSIT RIDERSHIP UP 10 PERCENT, SETS POST-PANDEMIC RECORD", Patrick Chandler
-   R Paper source code: https://github.com/thatonecalculator/math-28-project
